---
title: "rfcip (R FCIP)"

output:
  github_document:
    toc: true
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![R-CMD-check](https://github.com/dylan-turner25/rfcip/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dylan-turner25/rfcip/actions/workflows/R-CMD-check.yaml)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![codecov](https://codecov.io/gh/dylan-turner25/rfcip/graph/badge.svg?token=F6ZBMNEHYT)](https://codecov.io/gh/dylan-turner25/rfcip)
[![Coverage Status](https://coveralls.io/repos/github/dylan-turner25/rfcip/badge.svg)](https://coveralls.io/github/dylan-turner25/rfcip)

<!-- # Coverage Report -->

<!-- The `rfcip` integration with CodeCov is not currently working. Below are package coverage metrics as a substitute.  -->
<!-- ```{r, echo = F} -->
<!-- library(covr) -->
<!-- cov <- package_coverage(quiet = FALSE, clean = FALSE) -->
<!-- cov -->
<!-- ``` -->


# Introduction
`rfcip` provides a set of tools to allow users to access publicly available data related to the Federal Crop Insurance Program, including both Summary of Business data and Actuarial Data Master (ADM) datasets. The package provides a set of functions to easily navigate and access data that is publicly available, but otherwise scattered across different urls, files, and data portals. Although no official API exists for Federal Crop Insurance Data, much of the data can be located with a structured url meaning the `rfcip` package effectively functions like a defacto API wrapper.

**Disclaimer:** This product uses data provided by the USDA, but is not endorsed by or affiliated with USDA or the Federal Government.

# Why use the `rfcip` package?

Although the data that the `rfcip` package provides access to is all publicly available, it is often scattered across various files, interactive portals, or ftp sites and often comes in different file formats or arrives in files that are too large to work with in local memory. The `rfcip` package provides a set of functions that make it easy to access the data without having to think about these complexities. The package also makes it easy to update data for existing analysis since the provided functions can be easily integrated into a reproducible data pipeline for analysis or visualizations. This is particularly pertinent to the FCIP's primary data source, the Summary of Business, which is updated weekly. In short, the `rfcip` is a major time saver for anyone working with data related to the FCIP and has the added benefit of improving data accuracy and reproducibility. 


# Installation
`rfcip` can be installed directly from github using `remotes::install_github("https://github.com/dylan-turner25/rfcip")`

# Getting Started

Load the package and choose the years and filters you need. The year argument is
`year`, including when requesting several years.

```{r basic-usage, eval=FALSE}
library(rfcip)

# Summary of Business for corn in 2022
data <- get_sob_data(year = 2022, crop = "corn")

# Multiple years, with optional details about requests, retries, and sources
data <- get_sob_data(year = 2008:2020, log = TRUE)
```

By default, `get_sob_data()` uses a matching cached API report when available.
Otherwise, it tries RMA's interactive report API for each year. Temporary failures
such as HTTP 503 are retried up to four total attempts. If those attempts or the
retry wait budget are exhausted, it tries the coverage-level bulk files for that
year; the next year starts with the API again. Set `log = TRUE` to follow this
process. The default is `log = FALSE`; progress bars, warnings, and errors still
appear.

To download bulk files directly, without trying the interactive API:

```{r bulk-usage, eval=FALSE}
data <- get_sob_data(year = 2008:2020, sob_version = "sobcov")

# Detailed coverage-level records for corn in Iowa
corn <- get_sob_data(
  year = 2024, crop = "corn", state = "IA", sob_version = "sobcov"
)
```

Explicit `"sobcov"` calls return detailed state/county/crop/plan/coverage rows.
Automatic bulk backup instead matches the API query's grouping and compatible
columns. Bulk publication and program coverage can differ from the API, and
unavailable fields remain `NA`, so totals are not guaranteed to match. A warning
identifies each year using backup. Type/practice/unit detail remains available
through `sob_version = "sobtpu"`.

Repeated calls reuse available cached data. To request an update, use
`get_sob_data(year = 2024, crop = "corn", force = TRUE)`. A failed refresh can
return usable cached data with a warning; `force = TRUE` does not guarantee that
RMA will supply fresh data.

See the [getting-started vignette](vignettes/rfcip-introduction.md) for filters,
source limitations, cache management, and the package's other data sources, or
run `help(get_sob_data)` for the full function reference.
