---
title: "rfcip (R FCIP)"

output:
  github_document:
    toc: true
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE,
                      echo = TRUE,
                        fig.path = "man/figures/")
library(knitr)
library(dplyr)
library(devtools)
load_all()
```

[![R-CMD-check](https://github.com/dylan-turner25/rfcip/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dylan-turner25/rfcip/actions/workflows/R-CMD-check.yaml)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![codecov](https://codecov.io/gh/dylan-turner25/rfcip/graph/badge.svg?token=6sJUenko4X)](https://codecov.io/gh/dylan-turner25/rfcip)

## Introduction
`rfcip` provides a set of tools to allow users to access publicly available data related to the Federal Crop Insurance Program. The package provides a set of functions to easily navigate and access data that is publicly available, but otherwise scattered across different urls, files, and data portals. Although no official API exists for Federal Crop Insurance Data, much of the data can be located with a structured url meaning the `rfcip` package effectively functions like a defacto API wrapper.

**Disclaimer:** This product uses data provided by the USDA, but is not endorsed by or affiliated with USDA or the Federal Government.

## Installation
`rfcip` can be installed directly from github using `remotes::install_github("https://github.com/dylan-turner25/rfcip")`

## Supported Data Sources

### Summary of Business
The [Summary of Business](https://www.rma.usda.gov/tools-reports/summary-of-business) files produced by the USDA Risk Management Agency contains crop insurance participation measures and outcomes by state, county, crop, and insurance policy choices. Insured acres, collected premiums, disbursed subsides, liabilities, number of policies sold, number of indemnified policies, and loss ratios are all available from the summary of business. 

Accessing data from the summary of business can be done using the `get_sob_data`. With no arguments specified, the `get_sob_data` function will default to downloading data from [RMA's summary of business report generator](https://public-rma.fpac.usda.gov/apps/SummaryOfBusiness/ReportGenerato") for the current year, at the highest level aggregation. 

```{r sob-basic, include=T}
library(rfcip)
get_sob_data()
```

Most of the arguments for the `get_sob_data` function filter the returned data. For example, specifying the `year = 2022` and `crop = "corn"` will return data for corn in crop year 2022. For a description of all the arguments that can be supplied to `get_sob_data` see the help file for the function using `help(get_sob_data)`
```{r sob-filter, include=T}
get_sob_data(year = 2022, crop = "corn")
```
In the above example, the data set was filtered to a single crop, "corn". This potentially raises the question of what other values can be passed to the arguments that control the filters. For arguments with limited options, the options are explained in the functions help file (again, accessed via `help(get_sob_data)`). For some arguments that have many options, there are dedicated functions for pulling up values than can be passed to these arguments. 

The `get_crop_codes` function will return the commodity_code and commodity_name for a supplied year. Note that the `commodity_code` and `commodity_name` can be used interchangeably as values for the `crop` argument in `get_sob_data`. If there is a crop that is suspected to be an option, the `comm` argument can also be specified which will return just that specified crop if it exists. 
```{r crop-codes}
# get all crop codes and crop names for 2024
get_crop_codes(2024)
```
```{r crop-codes2}

# double check that "corn" is a valid crop
get_crop_codes(2024, crop = "corn")

# below are other ways to call the same data
get_crop_codes(2024, crop = "CORN")
get_crop_codes(2024, crop = 41)
```



The `get_insurance_plan_codes` function works analogously to the `get_crop_codes` function and helps identiy valid names of insurance plans. 
```{r get-ip}
# return all insurance plans avaliable in 2024
get_insurance_plan_codes(year = 2024)

# return the insurance plan code for the revenue projection plan
get_insurance_plan_codes(year = 2024, plan = "revenue protection")

# below are other ways to call the same data
get_insurance_plan_codes(year = 2024, plan = 2)
get_insurance_plan_codes(year = 2024, plan = "RP")
get_insurance_plan_codes(year = 2024, plan = "reVeNue PrOtEcTiOn")

```


As was previously stated above, most arguments for the `get_sob_data` function are for filtering the returned data. One exception is the `group_by` argument which does not filter the data being returned, but instead alters the level of aggregation. Taking the above example that returns data for corn in 2022 and setting `group_by = "county"` will return the same underlying as above, but decomposed by county. 
```{r sob-filter-group, include=T}
get_sob_data(year = 2022, crop = "corn", group_by = "county")
```

We can confirm `get_sob_data(year = 2022, crop = "corn")` and `get_sob_data(year = 2022, crop = "corn", group_by = "county")` return the same underlying data by summing up one of individual columns in the county level data. 
```{r sob-confirm, include=T}
national_data <- get_sob_data(year = 2022, crop = "corn")
print(paste("Liabilities from national data: ",sum(national_data$liabilities)))

county_data <- get_sob_data(year = 2022, crop = "corn", group_by = "county")
print(paste("Liabilities from county data:   ",sum(county_data$liabilities)))

```

A unique property of the summary of business data set is that its continuously updated (one per week) as new information is reported to USDA by [approved insurance providers](https://cropinsuranceinamerica.org/who-are-approved-insurance-providers-aips/). This means that analysis using the summary of business can quickly become outdated. One advantage of the the `rfcip` package is that it allows the raw data source to be directly integrated into the analysis. For example, the chart below plots indemnities for each crop year from 2015 up to the current year. The plot will automatically update with the latest data every time the plot is regenerated. Note that functions in `rfcip` are [memoised](https://en.wikipedia.org/wiki/Memoization) for the duration of the R session. This means that calling the same function with the same arguments will return a previously cached data set. In other words, in the below example, the data would not update if the code was run multiple times in the same R session, regaurdless of if the underlying data source changed. 

```{r sob-method-chaining, include=T}
library(rfcip)
library(dplyr)
library(ggplot2)

get_sob_data(year = 2015:as.numeric(format(Sys.Date(), "%Y"))) %>%
select(commodity_year, indemnity) %>%
group_by(commodity_year) %>%
summarize(indemnity = sum(indemnity)) %>%
mutate(indemnity = indemnity/1000000000) %>%
  ggplot(., aes(y = indemnity, x = commodity_year)) +
  geom_bar(stat = "identity", fill = "firebrick4") +
  xlab("") + ylab("USD (Billions)") +
  ggtitle(paste0("FCIP Indemnities by Year: 2015 - ",format(Sys.Date(), "%Y"))) +
  scale_y_continuous(labels = scales::dollar) +
  theme_minimal()

  

```


### Cause of Loss Files
Although RMA's [Summary of Business](https://www.rma.usda.gov/tools-reports/summary-of-business) files do report indemnities, they don't report the cause of loss associated with the indemnities which is often relevant. To obtain indemnities with the associated cause of loss that generated those indemnities, accessing the [cause of loss files](https://www.rma.usda.gov/tools-reports/summary-business/cause-loss) is necessary. Unlike data contained in the [summary of Business](https://www.rma.usda.gov/tools-reports/summary-of-business), the [cause of loss files](https://www.rma.usda.gov/tools-reports/summary-business/cause-loss) can only be accessed by bulk downloading all data for a particular year. Because of this, the cause of loss data cannot be filtered prior to loading it into local memory meaning there is no advantage to offering any within-function filtering options. To download [cause of loss files](https://www.rma.usda.gov/tools-reports/summary-business/cause-loss) for a year or series of years, use the `get_col_data` function. The function will automatically download all the relevant cause of loss files and merge them into a single data frame that will be returned. 


```{r col-basic, include=T}
col_data <- get_col_data(year = 2020:2022)
head(col_data)

```


### Price
Insurance guarantees and indemnities for revenue protection plans are base, in part, on projected commodity prices and harvest commodity prices. These are available via RMA's [price discovery application](https://public-rma.fpac.usda.gov/apps/PriceDiscovery/), but can also be obtained using the `get_price_data` function which supports arguments for `year`, `crop`, and `state`. For example, prices for corn in Illinois from 2020-2024 can be obtained with the following code. 

```{r price-basic, include=T}
price_data <- get_price_data(year = 2020:2024,
                             crop = "corn",
                             state = "IL")

head(price_data) 
```

### Reinsurance Agreements

#### Standard Reinsurance Agreement
Data related to the Standard Reinsurance Agreement (including retained liabilities, premiums, indemnities, and net underwriting gains and losses) is included as a static internal data set. To load the national level data set, use `data(nationalSRA)`. This data set is based on data from \url{https://www.rma.usda.gov/tools-reports/reinsurance-reports}. 
```{r national-sra, include=T}
# load the national SRA data set
data(nationalSRA)

head(nationalSRA)

# pull up the data sets documentation file.
?nationalSRA


```

#### Livestock Price Reinsurance Agreement
Data related to the Livestock Price Reinsurance Agreement (including retained liabilities, premiums, indemnities, and net underwriting gains and losses) is included as a static internal data set. To load the national level data set, use `data(nationalLPRA)`. This data set is based on data from \url{https://www.rma.usda.gov/tools-reports/reinsurance-reports}. 
```{r national-lpra, include=T}
# load the national SRA data set
data(nationalLPRA)

head(nationalLPRA)

# pull up the data sets documentation file.
?nationalLPRA

```



Please note that `rfcip` is released with a [Contributor Code of Conduct](https://ropensci.org/code-of-conduct/#:~:text=rOpenSci%20is%20committed%20to%20providing,understand%E2%80%9D%20or%20%E2%80%9CWhy%E2%80%9D.). By contributing to the package you agree to abide by its terms.