---
title: "rfcip (R FCIP)"

output:
  github_document:
    toc: true
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE,
                      echo = TRUE,
                        fig.path = "man/figures/")
library(knitr)
library(dplyr)
library(devtools)
load_all()
```

[![R-CMD-check](https://github.com/dylan-turner25/rfcip/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dylan-turner25/rfcip/actions/workflows/R-CMD-check.yaml)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![codecov](https://codecov.io/gh/dylan-turner25/rfcip/graph/badge.svg?token=F6ZBMNEHYT)](https://codecov.io/gh/dylan-turner25/rfcip)

# Coverage Report

The `rfcip` integration with CodeCov is not currently working. Below are package coverage metrics as a substitute. 
```{r, echo = F}
library(covr)
cov <- package_coverage(quiet = FALSE, clean = FALSE)
cov
```


# Introduction
`rfcip` provides a set of tools to allow users to access publicly available data related to the Federal Crop Insurance Program, including both Summary of Business data and Actuarial Data Master (ADM) datasets. The package provides a set of functions to easily navigate and access data that is publicly available, but otherwise scattered across different urls, files, and data portals. Although no official API exists for Federal Crop Insurance Data, much of the data can be located with a structured url meaning the `rfcip` package effectively functions like a defacto API wrapper.

**Disclaimer:** This product uses data provided by the USDA, but is not endorsed by or affiliated with USDA or the Federal Government.

# Installation
`rfcip` can be installed directly from github using `remotes::install_github("https://github.com/dylan-turner25/rfcip")`

# Getting Started
The `rfcip` package does not require any special setup or configuration. Once downloaded, data can be immediately accessed using any of the available functions. For example, the following pulls up RMA's summary of business data for corn in 2022. For a comprehensive introduction of the `rfcip` package, please see the [getting started](vignettes/rfcip-introduction.md) vignette.

```{r, basic-usage, include=T}

# get summary of business data for corn in 2022
library(rfcip)
get_sob_data(year = 2022, crop = "corn")


```

# Why use the `rcfip` package?

Although the data that the `rfcip` package provides access to is all publicly available, it is often scattered across various files, interactive portals, or ftp sites and often comes in different file formats or arrives in files that are too large to work with in local memory. The `rfcip` package provides a set of functions that make it easy to access the data without having to think about these complexities. The package also makes it easy to update data for existing analysis since the provided functions can be easily integrated into a reproducible data pipeline for analysis or visualizations. This is particularly pertinent to the FCIP's primary data source, the Summary of Business, which is updated weekly. In short, the `rfcip` is a major time saver for anyone working with data related to the FCIP and has the added benefit of improving data accuracy and reproducibility. 

