---
title: "Introduction to rfcip"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to rfcip}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(rfcip)
```

## Overview

The `rfcip` package provides tools to access publicly available data related to the Federal Crop Insurance Program (FCIP). The package functions as an unofficial API wrapper, making it easy to retrieve and analyze crop insurance data that is otherwise scattered across different URLs, files, and data portals.

**Important**: This package uses data provided by the USDA but is not endorsed by or affiliated with USDA or the Federal Government.

## Key Data Sources

### Summary of Business Data

The Summary of Business files contain comprehensive crop insurance participation and outcome data by state, county, crop, and insurance policy choices. This includes:

- Insured acres
- Collected premiums  
- Disbursed subsidies
- Liabilities
- Number of policies sold
- Number of indemnified policies
- Loss ratios

#### Basic Usage

Get current year data with default parameters:

```{r}
# Get current year summary data
current_data <- get_sob_data()
head(current_data)
```

#### Filtering Data

Filter by specific criteria:

```{r}
# Get corn data for 2022
corn_2022 <- get_sob_data(year = 2022, crop = "corn")

# Get data for multiple years and crops
multi_data <- get_sob_data(
  year = 2020:2022, 
  crop = c("corn", "soybeans"),
  state = "IA"
)
```

#### Discovering Available Options

Use helper functions to find valid parameter values:

```{r}
# Get all available crops for 2024
crops_2024 <- get_crop_codes(2024)
head(crops_2024)

# Get all available insurance plans
plans <- get_insurance_plan_codes(2024)
head(plans)

# Check if a specific crop exists
corn_check <- get_crop_codes(2024, crop = "corn")
```

### Detailed Data: Summary of Business by Type, Practice, and Unit Structure

For more granular data, use the SOBTPU version:

```{r}
# Get detailed corn data for 2022
detailed_data <- get_sob_data(
  year = 2022, 
  crop = "corn", 
  sob_version = "sobtpu"
)
head(detailed_data)
```

### Cause of Loss Data

Access indemnity data with associated causes of loss:

```{r}
# Get cause of loss data for multiple years
col_data <- get_col_data(year = 2020:2022)
head(col_data)
```

### Price Data

Retrieve commodity price data used for insurance calculations:

```{r}
# Get price data for corn in Illinois
price_data <- get_price_data(
  year = 2020:2024,
  crop = "corn", 
  state = "IL"
)
head(price_data)
```

### Livestock Data

Access livestock and dairy participation data:

```{r}
# Get Livestock Risk Protection data
lrp_data <- get_livestock_data(year = 2020:2022, program = "LRP")

# Get Dairy Revenue Protection data  
drp_data <- get_livestock_data(year = 2020:2022, program = "DRP")

# Get Livestock Gross Margin data
lgm_data <- get_livestock_data(year = 2020:2022, program = "LGM")
```

## Working with the Data

### Data Aggregation Levels

Control aggregation using the `group_by` parameter:

```{r}
# National level (default)
national <- get_sob_data(year = 2022, crop = "corn")

# County level detail
county_level <- get_sob_data(
  year = 2022, 
  crop = "corn", 
  group_by = "county"
)

# Multiple grouping variables
detailed <- get_sob_data(
  year = 2022:2024,
  group_by = c("insurance_plan", "cov_lvl")
)
```

### Exporting Data

Export data directly to Excel:

```{r}
# Export to Excel file (requires writexl package)
get_sob_data(
  year = 2022, 
  crop = "corn",
  dest_file = "corn_2022.xlsx"
)
```

## Performance Considerations

- Functions use memoization to cache results during R sessions
- For large datasets, consider filtering before downloading
- SOBTPU data requires full file downloads before filtering
- Progress bars show download status for long-running operations

## Static Datasets

The package includes pre-built datasets for reinsurance data:

```{r}
# National Standard Reinsurance Agreement data
data(nationalSRA)
head(nationalSRA)

# State-level SRA data
data(stateSRA)
head(stateSRA)

# National Livestock Price Reinsurance Agreement data
data(nationalLPRA)
head(nationalLPRA)
```

## Example Analysis

Here's a complete example analyzing corn insurance trends:

```{r}
library(dplyr)
library(ggplot2)

# Get corn data for recent years
corn_trends <- get_sob_data(
  year = 2015:2024,
  crop = "corn"
) %>%
  select(commodity_year, total_prem, indemnity, loss_ratio) %>%
  mutate(
    total_prem = total_prem / 1e9,  # Convert to billions
    indemnity = indemnity / 1e9
  )

# Plot premium vs indemnity trends
ggplot(corn_trends, aes(x = commodity_year)) +
  geom_line(aes(y = total_prem, color = "Premium")) +
  geom_line(aes(y = indemnity, color = "Indemnity")) +
  labs(
    title = "Corn Insurance: Premium vs Indemnity Trends",
    x = "Year",
    y = "Billions USD",
    color = "Measure"
  ) +
  theme_minimal()
```

## Getting Help

- Use `help(function_name)` for detailed function documentation
- Check function examples with `example(function_name)`
- Report issues on GitHub: https://github.com/dylan-turner25/rfcip
- All data comes directly from USDA RMA sources - see function documentation for source URLs